---
title: "Plotting Tape2Tape"
author: "Matthew Barlowe"
date: "3/28/2018"
output: html_document
---
## Tape 2 Tape Data

Ok today I'm going to be going over a few of the ways you can graph Tape 2 tape
data and show the different types of zone entrys and shot assists on a rink diagram.
For this excercise you will need R installed along with the `tidyverse` package.
`tidyverse` is actually a collection of packages, and you probably won't need 
all of them for this excercise, but I just find it's easier to load them all at 
once.

This tutorial will be written from the stand point of someone who is pretty new
to R so if you've just started don't worry. However, you have some familiarty 
with at least being able to run scripts in R, Rstuido, and some basic syntax 
such as `<-` assignment and function notation. Ok lets get to work.

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
```

## The Data

Ok first thing to do is to import the data into a dataframe. You can think of a
dataframe as kind of like an excel spreadsheet as there are rows and columns with
values located at the intersection of each row with each column. I'm going to
do this using the `read_csv` function that is in the `readr` package that will 
be loading into the script when we used `library()` on it up above.

`readr` is mainly a packge for reading in character delimited files such as csv
files but also tab delimited files, or my personal favorite `|` delimited files.
In this instance we will be reading in a csv file hence the use of `read_csv`.

The data comes in two files: one contains the actual play by play that has been 
tracked and another that has each player along with their skater ids. We'll need
that second file because as you'll see with the play by play file each player 
is identified by their id instead of their name. Ok lets get the data in and
get to work.

```{r cars}
file_name <- c('~/HockeyStuff/Tape2TapeData/Tape2Tape/11_10_plays.csv')
players_file_name <- c('~/HockeyStuff/Tape2TapeData/Tape2Tape/11_10_roster.csv')

pbp <- read_csv(file_name)
players <- read_csv(players_file_name)
```

Ok what I've done here is I've stored the path to the files as they would be on 
my computer into a variable and then read those variables into my dataframes I've
named `pbp` and `players`. You can just put the file path in the `read_csv` function
itself but is always good coding practice to store it in a variable and then use
that variable throughout the script. 

One reason is that it saves you from having 
to cut and paste/type a lot, and another is that if you wanted to run the same
script on another set of data all you would have to do is change that one variable
instead of multiple if you were reading it in at multiple points in your script. 

And so you can see the output that shows us that the files were read in succesfully.
You don't really need to know what the output is, but its just telling you that
it parsed certain columns with certain specifications such as the `periodTime` 
columns was read in as a time format etc. Again not much you need to worry about. 
Let's take a look at the data.

```{r pressure, echo=FALSE}
head(pbp)
```

The `head()` function will return the first six rows of any dataframe, and you 
can see all the different columns and we'll have a brief break down of each one:
```
period: the period the event took place in
periodTime: the score keepers time for each event
periodTimeRemaining: what the clock would be if you were watching the game
awayPlayer0Id...awayPlayer6Id: the six away players on the ice for event includes goalie
homePlayer0Id...homePlayer0Id: same as away but with home players
event: the main events tracked by Tape2Tape which include Zone Entry, Zone Exit, 
        Blocked Shot, Shot, Missed Shot, and Goal
eventType: this column further describes the event column. Will describe type of
            shot and type of Zone Exit/Entry as either Uncontrolled, Failed, or
            Controlled
eventResult: Lost or Recovered. This will only refer to Uncontrolled Zone Exits 
        and Entries
eventTeam: Tells the team that performed the event in the event column
x0, y0: This is the location of the event on the ice, there will only be a 
        corresponding x1, y1 value if the event was a zone exit created by a pass

plyer0Id, player1Id: Player0 is the player that performed the event in the event
                    column. If the event is a pass then player 0 is the passer and
                    player1 is the player that recieved the pass.
pass0x0, pass0y0...:These columns detail the passes that lead up to the event in 
                    the event column. pass0 is the pass right before the event,
                    pass1 is the pass before that etc.
```
The `players` dataframe is a lot more straightforward and the columns are 
self explanatory so we won't go over them. Ok let's convert all the user Ids
in the play by play dataframe to actual players names to make looking at the 
data and subsetting it easier.

```{r}
#convert player ids to actual player names from the player dataframe
convert_ids <- function(column, player_df){
    column <- player_df[match(column, player_df$playerId, nomatch = column), 
                        c('fullName')]
}

#converting playerids to playernames
pbp[4:15] <- pbp %>% select(awayPlayer0Id:homePlayer5Id) %>%
    sapply(convert_ids, player_df = players)

pbp[,c('player0Id', 'player1Id', 'pass0player0Id', 'pass0player1Id', 
       'pass1player0Id', 'pass1player1Id', 'pass2player0Id', 'pass2player1Id',
       'pass3player0Id', 'pass3player1Id', 
       'pass4player0Id', 'pass4player1Id')] <- 
    pbp[,c('player0Id', 'player1Id', 'pass0player0Id', 'pass0player1Id', 
           'pass1player1Id', 'pass1player1Id', 'pass2player0Id', 'pass2player1Id',
           'pass3player0Id', 'pass3player1Id', 'pass4player0Id', 
           'pass4player1Id')] %>% sapply(convert_ids, player_df = players)
```

Ok let's go over the code a bit here. The first part is a function I created 
that gets passed a column and a dataframe. This dataframe is the `players` dataframe
we created earlier. So this function goes along the column we pass to it and matches
the Id to the Id in the `players` dataframe and then returns the value from 
the `fullname` column in the `players` dataframe. 

Once it does that for 
every value in the column its returns those names and we store it in the same 
column variable and then we'll overwrite the the columns from the `pbp` 
dataframe with names instead of ids. 

The next part of the code is us applying
that function to all the columns we need to in the `pbp` dataframe using 
`sapply`. `sapply` is similar to a loop as it does a repeated action over
and over again given certain conditions, but behind the scenes things are quite
different. R is built around what's called vectorized operations because
R is slow when dealing with memory allocation and it is that reason that loops
run so slow in R. When you use the `apply` group functions or `plyr` functions
R automatically takes care of all those memory issues making things run faster
even though they actually use for loops inside their own source code.

But back to our `sapply` so all I'm doing is moving over all these columns and
looking up the player name for each player id and copying that to our new column.
This will allow us to break down our graphs to look at certain stats my player
name without having to consult the `players` dataframe anymore. So now that's 
done lets get to graphing.

## Creating the Rink
